\subsection{Problem 2}%
\label{sec:problem_2}
Perform the forward projection of the exact solution
$\begin{bmatrix} 1 & 0 & 1 & 1 & 0 \end{bmatrix}^T$
onto the range space spanned by the columns in the matrix:
\begin{equation*}
    A = \begin{bmatrix}
        1 & 2 & 2 & 3 & 1 \\
        2 & 4 & 4 & 6 & 2 \\
        3 & 6 & 6 & 9 & 6 \\
        1 & 2 & 4 & 5 & 3
\end{bmatrix}
\end{equation*}
Assuming the linear forward projection model, try to estimate the true solution to
$\matr{Ax} = \matr{b}$ given $\matr{A}$ and $\matr{b}$.
Then change the $a_{21}$ entry from 2 to 0 and repeat the estimation.
Explain the difference.
Compute the residual and solution errors.
Which algorithm gives the best estimate and why?
Which metrics are best to solve this problem?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Mathematics}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For an underdetermined system $\matr{Ax}=\matr{b}$ with
$\matr{A}\in\mathbb{R}^{m\times n}$, $m < n$, solutions exist if and only if
$\matr{b}\in\mathcal{R}(\matr{A})$ \citep{Meyer}.
When consistent, the solution set is the affine subspace
$\{\matr{x}_p + \matr{x}_h : \matr{x}_h \in \mathcal{N}(\matr{A})\}$,
which is infinite-dimensional whenever $\operatorname{rank}(\matr{A}) < n$.

\paragraph{Pseudoinverse.}
Among all solutions, the Moore--Penrose pseudoinverse selects the one of minimum
$\ell_2$ norm.
Via the singular value decomposition $\matr{A} = \matr{U\Sigma V}^T$ with
$r = \operatorname{rank}(\matr{A})$ non-zero singular values \citep{GoluVanl96}:
\begin{equation*}
  \matr{A}^+ = \matr{V}
  \begin{bmatrix} \matr{\Sigma}_r^{-1} & \matr{0} \\ \matr{0} & \matr{0} \end{bmatrix}
  \matr{U}^T.
\end{equation*}
The minimum-norm solution $\hat{\matr{x}} = \matr{A}^+\matr{b}$ lies in the row
space $\mathcal{R}(\matr{A}^T)$, orthogonal to $\mathcal{N}(\matr{A})$.

\paragraph{Effect of rank on exact recovery.}
The pseudoinverse recovers the true solution only if
$\matr{x}_{\mathrm{true}}\in\mathcal{R}(\matr{A}^T)$ \citep{Zdunek}.
For the base matrix with $\operatorname{rank}=3$, the null space has dimension
$5-3=2$; the true solution need not lie in the three-dimensional row space, so
exact recovery is not guaranteed.
Changing $a_{21}$ from 2 to 0 raises the rank to 4, shrinking the null space to
one dimension.
The true solution then lies in the four-dimensional row space and is the unique
minimum-norm solution, so the pseudoinverse recovers it to machine precision.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Solution}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
M-FOCUSS (Multiple-measurement FOCUSS) is a variant of the FOCUSS algorithm
that uses the same weight matrix $\matr{W}_k = \operatorname{diag}(|x_j^{(k)}|^{1-p/2})$
but applies it once rather than twice in the numerator of the update
\citep{GorodRao97}:
\begin{equation*}
  \matr{x}^{(k+1)}
  = \matr{W}_k \matr{A}^T
    \!\left(\matr{A}\matr{W}_k^2\matr{A}^T + \lambda\matr{I}\right)^{-1}
    \matr{b}.
\end{equation*}
Here we use $p=1$ and $\lambda=10^{-3}$.

The following MATLAB script computes $\matr{b}=\matr{Ax}_{\mathrm{true}}$, then
estimates the solution via pseudoinverse (\texttt{pinv}), FOCUSS
(Algorithm~\ref{algorithm:focuss}), and M-FOCUSS
(Algorithm~\ref{algorithm:mfocuss}) for both the base and modified matrices.

\lstinputlisting[style=Matlab-editor]{problems/Problem_2.m}

\paragraph{Residuals and solution errors.}
The residual $\|\matr{A\hat{x}}-\matr{b}\|_2$ is zero for any element of the
affine solution set, so it does not discriminate between estimates when the
system is rank-deficient.
The solution error $\|\hat{\matr{x}}-\matr{x}_{\mathrm{true}}\|_2$ is the
primary metric when the true solution is known.

\begin{center}
\begin{tabular}{llcc}
  \hline
  Matrix & Method & $\|\hat{\matr{x}}-\matr{x}_{\mathrm{true}}\|_2$ & $\|\matr{A\hat{x}}-\matr{b}\|_2$ \\
  \hline
  Base $\matr{A}$ (rank 3)     & Pseudoinverse & 0.90                & $8.7\times10^{-15}$ \\
                                & FOCUSS        & 1.73                & $8.1\times10^{-5}$  \\
                                & M-FOCUSS      & 1.08                & $3.7\times10^{-1}$  \\
  \hline
  Modified $\matr{A}$ (rank 4) & Pseudoinverse & $3.1\times10^{-15}$ & $5.7\times10^{-15}$ \\
                                & FOCUSS        & $1.7\times10^{-3}$  & $8.7\times10^{-4}$  \\
                                & M-FOCUSS      & $1.3\times10^{-2}$  & $8.5\times10^{-2}$  \\
  \hline
\end{tabular}
\end{center}

\paragraph{Best algorithm.}
For the base matrix ($\operatorname{rank}=3$), no algorithm recovers
$\matr{x}_{\mathrm{true}}$ exactly because the null space has dimension 2.
The pseudoinverse gives the smallest solution error ($0.90$) by returning the
minimum-$\ell_2$-norm consistent solution.
FOCUSS finds a sparser consistent solution $(0,0,0,2,0)^T$ but with larger error
($1.73$); M-FOCUSS does not fully converge to a consistent solution (residual
$0.37$).
For the modified matrix ($\operatorname{rank}=4$), $\matr{x}_{\mathrm{true}}$
lies in the row space and the pseudoinverse recovers it to machine precision.
FOCUSS and M-FOCUSS also converge close to the truth but less accurately due to
finite iterations and regularisation.

\paragraph{Best metrics.}
The residual $\|\matr{A\hat{x}}-\matr{b}\|_2$ is necessary but not sufficient:
it equals zero for the entire affine solution set when $\matr{b}\in\mathcal{R}(\matr{A})$.
When the true solution is known, solution error is the primary metric.
When it is not, the $\ell_2$ norm $\|\hat{\matr{x}}\|_2$ (minimised by the
pseudoinverse) or the $\ell_p$ quasi-norm $\|\hat{\matr{x}}\|_p^p$ (minimised
by FOCUSS/M-FOCUSS) serve as surrogates encoding prior knowledge about the
nature of the true solution.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
